A sheaf--function dictionary for algebraic tori over local fields

### 2013-10-03 shanghai.tex
Added a few more flags for our meeting this morning.

\begin{proof}
  Since $G$ is a smooth group scheme locally of finite type over $\Fq$, its
  identity component $G^0$ of $G$ is a connected smooth, 
  group scheme of finite type over $\Fq$, reduced over some finite extension of $\Fq$, by
   \cite{vdGeer-Moonen:AbelianVarieties}*{3.17}. 
	Since $\Fq$ is a finite field and hence perfect, $G^0$ is actually reduced over $\Fq$;
	see \cite{EGAIV2}*{Prop 6.4.1}.
	Since every group scheme over a field is separated 
	(see, for example, \cite{vdGeer-Moonen:AbelianVarieties}*{3.12}),  
	it follows that $G^0$ is a connected algebraic group.
    By Proposition~\ref{prop:connected}, every
  quasicharacter sheaf on $G^0$ is isomorphic to a 
  finite quasicharacter sheaf on $G^0$, 
  so to prove the proposition it suffices to show that every 
  finite quasicharacter sheaf on $G^0$ extends to a quasicharacter sheaf on $G$.

@@

  In the proof of Proposition~\ref{prop:restriction} we saw that $G^0$ is an algebraic group over $\Fq$. % clearly connected. 
  It now follows from Proposition~\ref{prop:connected} 
  that $\ker \trFrob{G^0} =0$ and $\coker \trFrob{G^0}=0$. 
  On the other hand, it follows from Propositions~\ref{prop:etale-iso} that $\ker \trFrob{\pi_0(G)}=0$
  and $\coker \trFrob{\pi_0(G)}=0$.

@@

  The sequence of groups of isomorphism classes of quasicharacter sheaves is right-exact: it is exact at
  $\QCiso{G^0}$ by Proposition~\ref{prop:restriction}; and
  it is exact at $\QCiso{G}$ because the left-square
  commutes (by Proposition~\ref{prop:pullback}), the bottom sequence
  is exact at $G(\Fq)^*$ and $\coker \trFrob{\pi_0(G)} =0$
  (by Proposition~\ref{prop:etale-iso}) and
  \[
  \image(\QCiso{\pi_0(G)} \to \QCiso{G}) \subseteq \ker(\QCiso{G} \to \QCiso{G^0}),
  \]
  by the last part of Proposition~\ref{prop:pullback}. Specifically,
  suppose $\qcs{L}$ is a quasicharacter sheaf on $G$ that is trivial on
  $G^0$. Then $\trFrob{\qcs{L}}\vert_{G^0(\Fq)} = \trFrob{\qcs{L}^0} =0$,
  by Proposition~\ref{prop:pullback} again. Since the sequence of
  character groups is exact at $G(\Fq)^*$, there is
  some $\chi \in \pi_0(G)(\Fq)^*$ such that the
  pull-back of $\chi$ along $G(\Fq)\to \pi_0(G)(\Fq)$ is
  $\trFrob{\qcs{L}}$. By Proposition~\ref{prop:etale-iso},
  there is a a quasicharacter sheaf $\qcs{L}_\chi$ on $\pi_0(G)$, unique up
  to isomorphism, such that $\trFrob{\qcs{L}_\chi} = \chi$. By
  Proposition~\ref{prop:pullback} again, $\qcs{L}$ is isomorphic to the
  image of $\qcs{L}_\chi$ under the monoidal functor $\QC(\pi_0(G)) \to \QC(G)$.

>>>>

\begin{proof}
\Clifton{Turn this paragraph into a lemma. We use it below.}  Since $G$ is a smooth group scheme locally of finite type over $\Fq$, its
  identity component $G^0$ of $G$ is a connected smooth, 
  group scheme of finite type over $\Fq$, reduced over some finite extension of $\Fq$, by
   \cite{vdGeer-Moonen:AbelianVarieties}*{3.17}. 
	Since $\Fq$ is a finite field and hence perfect, $G^0$ is actually reduced over $\Fq$;
	see \cite{EGAIV2}*{Prop 6.4.1}.
	Since every group scheme over a field is separated 
	(see, for example, \cite{vdGeer-Moonen:AbelianVarieties}*{3.12}),  
	it follows that $G^0$ is a connected algebraic group.

    By Proposition~\ref{prop:connected}, every
  quasicharacter sheaf on $G^0$ is isomorphic to a 
  finite quasicharacter sheaf on $G^0$, 
  so to prove the proposition it suffices to show that every 
  finite quasicharacter sheaf on $G^0$ extends to a quasicharacter sheaf on $G$.

@@

  \Clifton{Refer to the new lemma, here.}
  In the proof of Proposition~\ref{prop:restriction} we saw that $G^0$ is an algebraic group over $\Fq$. % clearly connected. 
  It now follows from Proposition~\ref{prop:connected} 
  that $\ker \trFrob{G^0} =0$ and $\coker \trFrob{G^0}=0$. 
  On the other hand, it follows from Propositions~\ref{prop:etale-iso} that $\ker \trFrob{\pi_0(G)}=0$
  and $\coker \trFrob{\pi_0(G)}=0$.

@@

  The sequence of groups of isomorphism classes of quasicharacter sheaves is right-exact:   it is exact at
  $\QCiso{G^0}$ by Proposition~\ref{prop:restriction}; and
  it is exact at $\QCiso{G}$ because the left-square
  commutes (by Proposition~\ref{prop:pullback}), the bottom sequence
  is exact at $G(\Fq)^*$ and $\coker \trFrob{\pi_0(G)} =0$
  (by Proposition~\ref{prop:etale-iso}) and   \Clifton{Awkward sentence.}
  \[
  \image(\QCiso{\pi_0(G)} \to \QCiso{G}) \subseteq \ker(\QCiso{G} \to \QCiso{G^0}),
  \]
  by the last part of Proposition~\ref{prop:pullback}. Specifically,
  suppose $\qcs{L}$ is a quasicharacter sheaf on $G$ that is trivial on
  $G^0$. Then $\trFrob{\qcs{L}}\vert_{G^0(\Fq)} = \trFrob{\qcs{L}^0} =0$,
  by Proposition~\ref{prop:pullback} again. Since the sequence of
  character groups is exact at $G(\Fq)^*$, there is
  some $\chi \in \pi_0(G)(\Fq)^*$ such that the
  pull-back of $\chi$ along $G(\Fq)\to \pi_0(G)(\Fq)$ is
  $\trFrob{\qcs{L}}$. By Proposition~\ref{prop:etale-iso},
  there is a quasicharacter sheaf $\qcs{L}_\chi$ on $\pi_0(G)$, unique up
  to isomorphism, such that $\trFrob{\qcs{L}_\chi} = \chi$. By
  Proposition~\ref{prop:pullback} again, $\qcs{L}$ is isomorphic to the
  image of $\qcs{L}_\chi$ under the monoidal functor $\QC(\pi_0(G)) \to \QC(G)$.
